---
title: 'Reproducible Research: Peer Assessment 1'
author: "Chris Hervochon"
date: "Sunday, November 15, 2015"
output: html_document
---

#Set the working directory
```{r}
require(knitr)
opts_knit$set(root.dir = 'c:/Users/Herv/My Documents')
```

#Load and preprocess the data
```{r}
df <- read.csv("activity.csv")
df$date <- as.Date(df$date)
```

Now we examine the mean total number of steps taken per day by means of a histogram
```{r}
library(ggplot2)
TotalStepsDay <- aggregate(x = df$steps , by = list(df$date), FUN = sum ,na.rm=TRUE)
names(TotalStepsDay) <- c("date","steps")
histplot <- ggplot(TotalStepsDay,aes(x = steps)) +
            ggtitle("Histogram of daily steps") +
            xlab("Steps (binwidth 2000)") +
            geom_histogram(binwidth = 2000)
histplot
```

The mean number of steps by day is calculated as:
```{r}
mean(TotalStepsDay$steps , na.rm = TRUE)
```

The median number of steps by day is calculated as:
```{r}
median(TotalStepsDay$steps , na.rm = TRUE)
```

Time series plot of 5-minute interval and the average number of steps taken, averaged across all days:
```{r}
AverageSteps  <- aggregate(x = df$steps , by = list(df$interval), FUN = mean ,na.rm=TRUE)
names(AverageSteps) <- c("interval","steps")

avg.step.line <- ggplot(AverageSteps,aes(interval,steps)) +
                 ggtitle("Time Series Plot of Average Steps by Interval") +
                 geom_line()
avg.step.line  

AverageSteps[which.max(AverageSteps$steps),c("interval")]
```

#Imputing missing values

```{r}
nrow(df[is.na(df$steps),])

df.imputed <- merge(x = df, y = AverageSteps, by = "interval", all.x = TRUE)
df.imputed[is.na(df.imputed$steps.x),c("steps.x")] <- df.imputed[is.na(df.imputed$steps.x),c("steps.y")]

df.imputed$date <- as.Date(df.imputed$date)
df.imputed$date.x <- NULL
df.imputed$Group.1 <- NULL
df.imputed$steps <- df.imputed$steps.x
df.imputed$steps.x <- NULL
df.imputed$steps.y <- NULL

TotalStepsDay <- aggregate(x = df.imputed$steps , by = list(df.imputed$date), FUN = sum ,na.rm=TRUE)
names(TotalStepsDay) <- c("date","steps")
histplot <- ggplot(TotalStepsDay,aes(x = steps)) +
            ggtitle("Histogram of daily steps after imputation") +
            xlab("Steps (binwidth 2000)") +
            geom_histogram(binwidth = 2000)
histplot 

mean(TotalStepsDay$steps , na.rm = TRUE)
median(TotalStepsDay$steps , na.rm = TRUE)
```

###Do these values differ from the estimates from the first part of the assignment? 
No, the values are the same.

###What is the impact of imputing missing data on the estimates of the total daily number of steps?
There is very little, if any, impact.

###Are there differences in activity patterns between weekdays and weekends?
Yes, the patterns are different, per the below.
```{r}
df.imputed$weekday <- as.factor(ifelse(weekdays(df.imputed$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday")) 

AverageSteps.and.weekday  <- aggregate(x = df.imputed$steps , 
                                                    by = list(df.imputed$interval,df.imputed$weekday), FUN = mean ,na.rm=TRUE)
names(AverageSteps.and.weekday) <- c("interval","weekday","steps")

avg.step.line <- ggplot(AverageSteps.and.weekday,aes(interval,steps)) +
                 ggtitle("Time Series Plot of Average Steps by Interval after Imputation") +
                 facet_grid(. ~ weekday) +
                 geom_line(size = 1)
avg.step.line  
```
